# Transformations

This chapter contains the data transformations we performed before analyzing the data.

## Data Type Conversion

Before we analyze the data, we need to convert some of the columns.

```{r}
#| code-fold: false
library(tidyverse)

experiments <- read_csv("data/experiments.csv")

algorithm_levels <- c("sobol", "gpei", "turbo", "cma-es")
algorithm_labels <- c("Sobol", "GPEI", "TurBO", "CMA-ES")

experiments <- 
  experiments |> 
  mutate(bom_id = factor(bom_id),
         replication = as.integer(replication),
         budget = as_factor(budget),
         candidate_id = as.integer(candidate_id),
         num_trials = as.integer(num_trials),
         num_candidates = as.integer(num_candidates),
         algorithm = factor(algorithm, levels = algorithm_levels, labels = algorithm_labels)
         )

experiments 
```

We will load the transformed data from an R-data source (RDS) file in the following chapters:

```{r}
#| code-fold: false
experiments |> 
  saveRDS("data/experiments.rds")
```


## Consistent Colors and Theming

Define a consistent color set for the display of the `algorithm` variable throughout the analysis:

```{r}
#| code-fold: false
# Consistent colors
library(ggsci)

algorithm_colors <- c(
  "Sobol" = pal_aaas()(4)[1],
  "GPEI" = pal_aaas()(4)[2],
  "TurBO" = pal_aaas()(4)[3],
  "CMA-ES" = pal_aaas()(4)[4]
)
```

Define the globally used theme:

```{r}
#| code-fold: false
theme_simbo <- function (base_size = 12, base_family = "Roboto Condensed") {
  half_line <- base_size/2
  theme_bw()
}

theme_set(theme_simbo())
```

Example:

```{r}
library(scales)

experiments |> 
  ggplot() +
  aes(x = algorithm, fill = algorithm) +
  geom_bar() +
  scale_fill_manual(values = algorithm_colors) +
  scale_y_continuous(labels = comma) +
  labs(x = "Algorithm", y = "Number of Candidates", fill = "Algorithm")
```


